\tool: Programmable, Efficient, and Scalable Secure Two-Party Computation

### 2017-10-27 OaklandSubmission/intro.tex
some edits in the intro

  \TODO{While taking a pass on intro, check citation correctness or if we are missing any citation.}

@@

Today it is hard for the developers to program secure applications
using cryptographic techniques. Typical developers lack a deep
understanding of the cryptographic protocols, and therefore cannot be
expected to use them correctly and efficiently on their own.
Ideally, a developer would declare the functionality in a general
purpose, high-level programming language and a
tool, e.g. a compiler, would generate an efficient protocol that
implements the functionality securely, while hiding the cryptography
behind-the-scenes. This paper presents such a framework for Secure
Two-party Computation (\mpc).

@@

\mpc~(\cite{yao,gmw}) is a powerful
cryptographic technique that allows two mutually distrusting parties
to compute a joint function of their secret inputs in a way that both
the parties learn nothing about the inputs of each other beyond what
is revealed by their (possibly different) outputs. For example, \mpc
can be used for {\em secure
prediction}~(\cite{shafindss,wu,barni,minionn,secureml}),
where one party (the server) holds a proprietary classifier to predict
a label (e.g. a disease, genomics, or spam detection), and the other
party (the client) holds a private input that it wants to run the
classifier on. Using \mpc guarantees that the server learns nothing
about the client's input or output, and that the client learns nothing
about the server model, beyond what is revelaed by the output
label.

Since its introduction in 1980s, there has been a long sequence
of work that has transformed \mpc from a mere theoretical concept to
something practically efficient and usable. To understand the
state-of-the-art, let us consider the following example used by many
secure prediction algorithms. Suppose Alice wants to write
a \mpc protocol to securely
compute $w^Tx >b$. Here $w$ and $x$ are two vectors (inputs of the two
parties), $b$ is a scalar, $\cdot^{T}$ is the matrix transpose
operator, and $w^Tx$ denotes the inner product of
$w^{T}$ and $x$.  With the current state-of-the-art, Alice has following
options:

@@

She can program the computation in one of the several programmer
friendly, domain-specific languages (such
as Fairplay~\cite{fairplay}, Wysteria~\cite{wysteria},
ObliVM~\cite{oblivm}, CBMC-GC~\cite{cbmcgc}, SMCL~\cite{smcl},
Sharemind~\cite{sharemind}, \cite{lambdaps} etc.) that would
automatically compile it to a \mpc protocol. However, all of these
frameworks use cryptographic backends that take as input
the computation expressed as either a boolean
circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{gentry}).
%
The efficiency of the generated \mpc protocol is thus bounded by the
efficiency of representing the computation in one of
these representations. For instance, multiplication of two
$\ell$-bit integers can be either expressed as a boolean circuit of
size $O(\ell^2)$, or as an arithmetic circuit with 1 multiplication
gate.
%
Hence, for better efficiency, Alice would ideally like to compute
$w^Tx$ using an arithmetic circuit, and the comparison with $b$ using
a boolean circuit.
%
Unfortunately, none of the above frameworks support combinations of
arithmetic and boolean circuits, and using different tools for
different parts of the computation is cumbersome and error-prone.

@@

This paper presents \tool\footnote{Read as ``easy peasy'',
stands for Easy 2 Party Computation}, a new \mpc framework.
\tool is the first
``cryptographic-cost aware'' framework that generates efficient and
scalable \mpc protocols, comprising combinations of arithmetic and
boolean circuits, from high-level programs devoid of any
cryptography. Below we describe the salient features of \tool.

\subsubsection*{Programmability} \tool source programs are ideal
functionalities that describe ``what'' computation needs to be done,
rather than ``how'' to do it. In particular, the programmer writes the
high-level computation without thinking about the underlying
cryptographic details. For example, an \tool source
program for $w^Tx >b$ is as follows:

@@

\vspace{0.2cm}
We note that the program is quite similar to what a programmer might
write in a common programming language. The simplicity of the
language comes with its usual benefits: it is easily accessible to the
developers, there are fewer avenues for making mistakes, and it is
easy to maintain and modify the programs. Needless to say, frameworks
that expose low-level circuit APIs to the programmer do not enjoy
these benefits.

\subsubsection*{Cryptogaphic-cost awareness} \mpc compiler is aware
of the cryptographic costs of arithmetic and boolean representations
of the source language operators, and their corresponding
inter-conversion gates. Based on these costs, the compiler, given a
source program, automatically picks arithmetic or boolean
representations for different parts of the program, and generates the
corresponding circuits along with the required inter-conversion
gates. The outcome is an effcient \mpc protocol using a combination of
arithmetic and boolean circuits, while the programmer remains
oblivious of all the low-level cryptographic details. Our evaluation
shows that the performance of the
protocols automatically generated by the \tool compiler is comparable
to or better than the custom, specialized protocols developed in the
previous work. Indeed, \tool is the first such cryptographic-cost
aware compiler.

\subsubsection*{Scalability (secure code partitioning)} \mpc tools
often do not scale to large programs. The reason is that
most \mpc implementations use a circuit-like representation as an
intermediate language. Hence, the largest compute that can be done
securely is upper-bounded by the largest circuit that can fit in the
memory on a machine\footnote{Using swap and
disk space is feasible but it causes huge slowdown.}. This is a
show-stopper for applications like secure machine learning, secure
prediction, etc. that operate on large amounts of data. \aseem{Can we
cite something here to support this claim?}
\tool addresses the scalability concern using a novel technique that
we call secure code partitioning (or partitioning in short). At
a high level, we break the original program into a sequence of small
sub-programs, which are then sequentially processed by our
framework, while appropriately threading the intermediate outputs
along. While this
addresses the scalability concern (the circuit
sizes of the sub-programs are now small enough to fit in the memory),
we still have to address
the security risk of revealing the intermediate outputs. To address
this, \tool (a) masks the intermediate outputs using a secure secret
sharing between the two parties, and (b) appends the
share-reconstruction function to the beginning of the subsequent
sub-program. As we show in our evaluation, partitioning allows us to
program large applications in \tool.

@@

\subsubsection*{Formal guarantees} We prove a formal correctness and
security theorem for our compiler. The correctness theorem relates the
``ideal'' semantics (the trusted-third party semantics) of a source
program and the ``protocol'' semantics (the distributed \mpc
semantics that relies on circuit evaluation) of the corresponding
compiled program. Both these semantics are instrumented
with \emph{observations} that record the outputs. We prove that a
well-typed source program terminates successfully in the source
semantics. In addition, we prove that for a well-typed source program,
its observations in the source semantics are same as the observations
of its corresponding compiled program in the protocol
semantics. {\aseem{Had a hard time writing this last sentence,
suggestions welcome.}}
For the security theorem, we formally reduce the security of our
compiler to the security of the back-end cryptographic
protocol.

@@

{\aseem{Not sure about this baseline paragraph. Did not touch it.}} 
\subsubsection*{Performance Baseline} Most works cite garbled
circuit based two-party computation as the most practical approach so
far and use it as the baseline for comparison of performance. But, it
has been shown multiple times that the garbled circuits based approach
does not scale for realistic examples. Even for small benchmarks
(e.g., Na\"{i}ve Bayes) the size of garbled circuit can be
huge. Hence, comparisons are often done on artificial data sizes and
garbled circuits are shown to be 500x slower than hand-crafted
protocols \cite{shafindss}. This is cited as the motivation for need
for specialized protocols for different ML algorithms in secure
prediction such as linear regression \cite{shafindss}, decision
trees \cite{wu}, deep neural networks \cite{minionn} and matrix
factorization \cite{valeriaMatrix}. Our work is based on generic \mpc
and we show competitive performance with the specialized protocols
contrary to the popular belief. Hence, we believe \tool based
implementations can serve as the performance baseline for future works
on specialized protocols.

@@

\aseem{Checkpointing here. Two more paragraphs to come, one for
summary of implementation and evaluation, and one for related work.}

>>>>

  %% \TODO{While taking a pass on intro, check citation correctness or if we are missing any citation.}

@@

%% \aseem{The intuition of this paragraph is conveyed in rest of the
%% intro already?}

%% Today it is hard for the developers to program secure applications
%% using cryptographic techniques. Typical developers lack a deep
%% understanding of the cryptographic protocols, and therefore cannot be
%% expected to use them correctly and efficiently on their own.
%% Ideally, a developer would declare the functionality in a general
%% purpose, high-level programming language and a
%% tool, e.g. a compiler, would generate an efficient protocol that
%% implements the functionality securely, while hiding the cryptography
%% behind-the-scenes. This paper presents such a framework for Secure
%% Two-party Computation (\mpc).

@@

Secure Two-party Computation (\mpc)~(\cite{yao,gmw}) is a powerful
cryptographic technique that allows two mutually distrusting parties
to compute a joint function of their secret inputs in a way that both
the parties learn nothing about the inputs of each other beyond what
is revealed by their (possibly different) outputs. For example, \mpc
can be used for {\em secure
prediction}~(\cite{shafindss,wu,barni,minionn,secureml}),
where one party (the server) holds a proprietary classifier to predict
a label (e.g. a disease, genomics, or spam detection), and the other
party (the client) holds a private input that it wants to run the
classifier on. Using \mpc guarantees that the server learns nothing
about the client's input or output, and that the client learns nothing
about the server model, beyond what is revelaed by the output
label.

Since its introduction in 1980s, there has been a long sequence
of work that has transformed \mpc from a mere theoretical concept to
something practically efficient and usable. To understand the
state-of-the-art, let us consider the following example underlying many
secure prediction algorithms. Suppose Alice wants to write
a \mpc protocol to securely
compute $w^Tx >b$. Here $w$ (a vector) and $b$ (a scalar) constitute
the server model, and $x$ is the client's input vector. Further,
$\cdot^{T}$ is the matrix transpose operator, and $w^Tx$ denotes the
inner product of  $w^{T}$ and $x$. With the current state-of-the-art,
Alice has following options:

@@

She can program the computation in one of the several programmer
friendly, domain-specific languages (such
as Fairplay~\cite{fairplay}, Wysteria~\cite{wysteria},
ObliVM~\cite{oblivm}, CBMC-GC~\cite{cbmcgc}, SMCL~\cite{smcl},
Sharemind~\cite{sharemind}, \cite{lambdaps} etc.) that would
automatically compile it to a \mpc protocol. However, all of these
frameworks use cryptographic backends that take as input
the computation expressed as either a boolean
circuit~(\cite{yao,gmw}) or an arithmetic circuit~(\cite{gentry}).
%
The efficiency of the generated \mpc protocol is thus bounded by the
efficiency of representing the computation in \emph{one} of
these representations. For instance, multiplication of two
$\ell$-bit integers can be either expressed as a boolean circuit of
size $O(\ell^2)$, or as an arithmetic circuit with 1 multiplication
gate.
%
Hence, for better efficiency, Alice would ideally like to compute
$w^Tx$ using an arithmetic circuit, and the comparison with $b$ using
a boolean circuit.
%
Unfortunately, none of the above frameworks support combinations of
arithmetic and boolean circuits, and using different tools for
different parts of the computation is cumbersome and error-prone.

@@

This paper presents \tool\footnote{Read as ``easy peasy'',
stands for Easy 2 Party Computation}, a new \mpc framework.
\tool is the first
``cryptographic-cost aware'' framework that generates efficient and
scalable \mpc protocols from high-level programs devoid of any
cryptographic details. The generated protocols comprise combinations
of arithmetic and boolean circuits and have performance comparable to
or better than the custom, specialized
protocols from previous work. Below we describe the
salient features of \tool.

\subsubsection*{Ease of programming} \tool source programs are ideal
functionalities that describe ``what'' computation needs to be done,
rather than ``how'' to do it. In particular, the programmer writes the
high-level computation without thinking about the underlying
cryptographic details. For example, an \tool source
program for $w^Tx >b$ is as follows:

@@

\vspace{0.2cm}
The program is quite similar to what a programmer might
write in C++ or Java. The simplicity of the
language comes with its usual benefits: it is easily accessible to the
developers, there are fewer avenues for making mistakes, developers
don't bear the burden of getting cryptographic details right, and it
is easy to maintain and modify the programs. Needless to say,
frameworks that expose low-level circuit APIs to the programmer do not
enjoy these benefits.

\subsubsection*{Cryptogaphic-cost aware compiler} The \tool compiler
compiles a source program to a hybrid computation consisting
of \emph{public} and \emph{secret} parts. In the example above, for
instance, \tool compiler realizes that the array index~\ls{i} is
public, and hence, the array accesses need not be compiled as
multiplexers. Further, within the secret parts, \tool compiler is
aware
of the cryptographic costs of arithmetic and boolean representations
of the source language operators. Based on these costs, the compiler
automatically picks arithmetic or boolean
representations for different sub-parts, and generates the
corresponding circuits along with the required inter-conversion
gates. The outcome is an effcient \mpc protocol combining arithmetic
and boolean circuits, while the programmer remains
oblivious of all these cryptographic details. Our evaluation
shows that the performance of these automatically generated protocols
is comparable
to or better than the custom, specialized protocols developed in the
previous work. Indeed, \tool is the first such cryptographic-cost
aware compiler.

\subsubsection*{Scalability (secure code partitioning)} \mpc tools
often do not scale to large programs. The reason is that
most \mpc implementations use a circuit-like representation as an
intermediate language. Hence, the largest compute that can be done
securely is upper-bounded by the largest circuit that can fit in the
machine memory\footnote{Using swap and
disk space is feasible but it causes huge slowdown.}. This is a
show-stopper for applications like secure machine learning, secure
prediction, etc. that operate on large amounts of data. \aseem{Can we
cite something here to support this claim?}
\tool addresses the scalability concern using a novel technique that
we call secure code partitioning (or partitioning in short). At
a high level, we break the original program into a sequence of small
sub-programs, which are then sequentially processed by \tool, while
appropriately threading the intermediate outputs
along. While this
addresses the scalability concern (the circuit
sizes of the sub-programs are now small enough to fit in the memory),
we still have to address
the security risk of revealing the intermediate outputs. To address
this, \tool (a) masks the intermediate outputs using a secure secret
sharing between the two parties, and (b) appends the
share-reconstruction function to the beginning of the subsequent
sub-program. As we show in our evaluation, partitioning allows us to
program large applications in \tool.

@@

\subsubsection*{Formal guarantees} We prove formal correctness and
security theorems for our compiler. The correctness theorem proves a
forward simluation between the
``ideal'' semantics (the trusted-third party semantics) of a source
program and the ``protocol'' semantics (the distributed \mpc
semantics that relies on circuit evaluation) of the corresponding
compiled program. The simulation
theorem not only guarantees that a
well-typed source program terminates successfully in the ideal
semantics (e.g. there are no array index out-of-bounds errors), but
also that the observable outputs in the two semantics are the same. For
the security theorem, we formally reduce the security
of our compiler to the security of the back-end cryptographic
protocol. We also prove a formal security theorem for our partitioning
scheme.

@@

%% {\aseem{Not sure about this baseline paragraph. Did not touch it.}} 
%% \subsubsection*{Performance Baseline} Most works cite garbled
%% circuit based two-party computation as the most practical approach so
%% far and use it as the baseline for comparison of performance. But, it
%% has been shown multiple times that the garbled circuits based approach
%% does not scale for realistic examples. Even for small benchmarks
%% (e.g., Na\"{i}ve Bayes) the size of garbled circuit can be
%% huge. Hence, comparisons are often done on artificial data sizes and
%% garbled circuits are shown to be 500x slower than hand-crafted
%% protocols \cite{shafindss}. This is cited as the motivation for need
%% for specialized protocols for different ML algorithms in secure
%% prediction such as linear regression \cite{shafindss}, decision
%% trees \cite{wu}, deep neural networks \cite{minionn} and matrix
%% factorization \cite{valeriaMatrix}. Our work is based on generic \mpc
%% and we show competitive performance with the specialized protocols
%% contrary to the popular belief. Hence, we believe \tool based
%% implementations can serve as the performance baseline for future works
%% on specialized protocols.

@@

\aseem{Checkpointing here. Three more paragraphs to come, one for
summary of implementation and evaluation, one for related work, and
one for baseline argument.}
